---
title: "Sources of Real Economic Growth - PDF and LaTeX"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gt)
library(tidyverse)
library(readxl)
```

See discussion on the [RStudio Community Table Gallery](https://community.rstudio.com/t/simple-latex-table-example-sources-of-real-economic-growth/136408)

This is adapted from this [LinkedIn Post]( https://www.linkedin.com/feed/update/urn:li:activity:6916876075809558528/), by [Mike Jadoo](https://www.linkedin.com/in/mike-jadoo/).


# Table - Sources of Real Economic Growth

```{r}

gt_tbl <-
  readxl::read_excel("sources_of_real_economic_growth.xlsx", sheet = "data") %>%
  dplyr::select(Years, Output, LaborC, CapitalC, TFP) %>%
  gt() %>%

  #center align columns
  cols_align(align = c("center"), columns = everything()) %>%

  # Add column labels
  cols_label(
    Years = "Periods",
    Output = "Real GDP",
    LaborC = md("Labor Input • (*a*)"),
    CapitalC = md("Capital Input • (1 - *a*)"),
    TFP = "Total Factor Productivity"
  ) %>%

  # resize Capital Input column
  cols_width(CapitalC ~ pct(20)) %>%

  # resize font size.
  tab_options(
    table.width = pct(80),
    table.font.size = "smaller",
    column_labels.font.size = "small"
  ) %>%

  # Add table title and subtitle
  tab_header(
    title = "Source of Real Economic Growth",
    subtitle = "Private Non-Farm Business"
  ) %>%

  # Add and title a column label spanner
  tab_spanner(
    label = html("= Labor Input* + Capital Input* + TFP"),
    columns = matches("LaborC|CapitalC|TFP")
  ) %>%

  # Add and format footnotes
  # Note the md() function to handle markdown, will convert this to HTML,
  # LaTeX or other, defined by yaml output parameter.
  tab_source_note(md("\\* Contribution points towards output growth.")) %>%
  tab_source_note(md("Here, *a* is the average labor cost share.")) %>%
  tab_source_note(md("Source: Bureau of Labor Statistics.")) %>%
  tab_footnote(
    footnote = "Excludes Nonprofits, Private Households, Owner-occupied Housing,
    Government and Government Enterprise.",
    locations = cells_column_labels(columns = Output)
  ) %>%
  tab_options(source_notes.multiline = FALSE)

gt_tbl

```


## LaTeX Output

In the code chunk above we created out `gt` table. Suppose we want to export this into a format we can put into a LaTeX document, how would we approach that? 

The [`as_latex`](https://gt.rstudio.com/reference/as_latex.html) function takes a `gt()` table object as input, and returns a `knit_asis` object which an R Markdown document, which makes it easy to include in R Markdown documents that are knited to PDF.

If you'd just like the LaTeX code, for example to paste that into a separate LaTeX document, use `cat()`, see below. Use `as.character()` instead of `cat` if you'd like this LaTeX table code in a single line. 

```{r comment = ""}

gt_tbl %>% as_latex() %>% cat() 

```